Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-02-06 paper.qmd
Changed the text of effect density

To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $7\text{-}D$ data having C-shaped structure is used. @fig-one-dens_clust-error a1 and b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used. The \gD{} layouts are colored according to the high-dimensional model error. In b1, the large model error points are clustered in the top left corner, forming a separate cluster with high model errors.

The models are fitted to these layouts. @fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space. In these projections, the differences in the distribution of \pD{} model errors can be observed. The uniform C-shaped structure displays randomly distributed model errors, while the dense C-shaped structure shows high model errors at the sparse end. In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.

>>>>

To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $4\text{-}D$ data having C-shaped structure is used. @fig-one-dens_clust-error b shows \gD{} layouts for tSNE. The default hyper-parameters are used. The \gD{} layout is colored according to the $4\text{-}D$ model error (@fig-one-dens_clust-error a). In @fig-one-dens_clust-error a, the large model error points are clustered in the top corner, forming a separate cluster with high model errors.

The model is fitted to this layout. @fig-one-dens_clust-error c shows the $2\text{-}D$ projection of the fitted model. In this projection, the distribution of $4\text{-}D$ model errors can be observed. The dense C-shaped structure shows high model errors at the sparse end. In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
